---
title: "ST 558 Project 2 Report"
author: "Collin Knezevich & Jiyue Zhang"
date: '2022-07-02'
output: html_document
---

```{r}
# Libraries 
library(tidyverse)
```

# Introduction 

# Data 

We will read in the online news popularity data set locally. 
```{r}
newsData <- read_csv("OnlineNewsPopularity/OnlineNewsPopularity.csv")
```

For our initial analysis, we will only consider data from a single data channel. We will consider the "bus" (business) data channel, and we will subset the data to focus on this channel. 
```{r}
busNews <- newsData %>% filter(data_channel_is_bus == 1) 
```

# Summarizations 

Our response variable of interest for this analysis is the number of shares. First, we will quickly summarize the distribution of the shares variable. 
```{r}
summary(busNews$shares)
```
Here, we can see the mean number of shares, as well as the five number summary (min, 1st quartile, median, 3rd quartile, max).   

Next, we would like to compare the proportion of articles that are shared a greater number of times than the median number of shares, for articles that are published on a weekend day or not. We will create a new variable, equal to 1 if the article was shared a greater number of times than the median, and equal to 0 otherwise. 
```{r}
md <- median(busNews$shares)
busNews2 <- busNews %>% mutate(shares_gt_median = ifelse(shares > md, 1, 0)) 

ggplot(data = busNews2, aes(x = as.factor(is_weekend))) + 
  geom_bar(aes(fill = as.factor(shares_gt_median)), position = "fill") + 
  labs(title = "Proportion of Articles Shared More Than the Median, 
       for Published on Weekend vs. Not", x = "Weekend", y = "Proportion", 
       fill = "Shares > Median") + 
  scale_x_discrete(labels = c('No', 'Yes')) + 
  scale_fill_discrete(labels = c('No', 'Yes'))
```
We can easily compare the difference between articles published on the weekend and articles published on weekdays. 
